---
title: "Iowa Legislature - 91st General Assembly"
format:
  html:
    echo: false
    message: false
    warning: false
    toc: true
    toc-location: right
    toc-depth: 2
---

```{r}
library(readr)
library(dplyr)
library(stringr)
library(gt)
library(ggplot2)
library(ggparliament)
source(here::here("R/utils.R"))
```

```{r}
# Load data
bills    <- read_csv(here::here("legiscan/files_ga91_derived/bills.csv"),     show_col_types = FALSE)
history  <- read_csv(here::here("legiscan/files_ga91_derived/history.csv"),   show_col_types = FALSE)
people   <- read_csv(here::here("legiscan/files_ga91_derived/people.csv"),    show_col_types = FALSE) |>
  filter(!is.na(first_name), first_name != "")
votes     <- read_csv(here::here("legiscan/files_ga91_derived/votes.csv"),    show_col_types = FALSE)
rollcalls <- read_csv(here::here("legiscan/files_ga91_derived/rollcalls.csv"),show_col_types = FALSE)

bill_groups_path <- here::here("data/bills/groups.rds")
if (file.exists(bill_groups_path)) {
  bill_groups <- read_rds(bill_groups_path)
} else {
  bill_groups <- NULL
}
```

```{r}
# Identify replaced legislators (same logic as legislators/index.qmd)
last_vote_dates <- votes |>
  left_join(rollcalls |> select(roll_call_id, date), by = "roll_call_id") |>
  group_by(people_id) |>
  summarize(last_vote_date = max(date, na.rm = TRUE), .groups = "drop")

dup_districts <- people |>
  count(role, district) |>
  filter(n > 1)

former_ids <- people |>
  semi_join(dup_districts, by = c("role", "district")) |>
  left_join(last_vote_dates, by = "people_id") |>
  group_by(role, district) |>
  mutate(
    is_former = case_when(
      all(!is.na(last_vote_date)) ~ last_vote_date == min(last_vote_date),
      !is.na(last_vote_date)     ~ TRUE,
      is.na(last_vote_date)      ~ FALSE
    )
  ) |>
  ungroup() |>
  filter(is_former) |>
  pull(people_id)

active_people <- people |> filter(!people_id %in% former_ids)

# Party composition for each chamber
party_colours <- c("R" = "#C0392B", "D" = "#232066")

house_comp  <- active_people |> filter(role == "Rep") |> count(party, name = "seats") |> arrange(desc(seats))
senate_comp <- active_people |> filter(role == "Sen") |> count(party, name = "seats") |> arrange(desc(seats))

n_house_R  <- house_comp  |> filter(party == "R") |> pull(seats)
n_house_D  <- house_comp  |> filter(party == "D") |> pull(seats)
n_senate_R <- senate_comp |> filter(party == "R") |> pull(seats)
n_senate_D <- senate_comp |> filter(party == "D") |> pull(seats)
```

Welcome to the Iowa Legislature tracker for the **91st General Assembly (2025)**. This site provides access to bill information, legislator profiles, voting records, and lobbyist declarations.

## Quick Stats

```{r}
n_bills <- nrow(bills)
n_house_bills <- sum(str_detect(bills$bill_number, "^HF|^HJR|^HR"))
n_senate_bills <- sum(str_detect(bills$bill_number, "^SF|^SJR|^SR"))
```

::: {.grid}

::: {.g-col-6}
**Legislation**

- **`r format(n_bills, big.mark=",")`** total bills
- `r format(n_house_bills, big.mark=",")` House bills
- `r format(n_senate_bills, big.mark=",")` Senate bills
:::

::: {.g-col-6}
**Bill Progress**

```{r}
#| results: asis
status_order <- c("Introduced", "Passed Subcommittee", "Passed Committee",
                  "Passed One Chamber", "Passed Both Chambers", "Signed by Governor")

if (!is.null(bill_groups)) {
  stage_counts <- bill_groups |>
    count(status, name = "n") |>
    mutate(status = factor(status, levels = status_order)) |>
    arrange(desc(status))

  for (i in seq_len(nrow(stage_counts))) {
    cat(sprintf("- **%s** %s\n", format(stage_counts$n[i], big.mark = ","),
                as.character(stage_counts$status[i])))
  }
} else {
  cat("_Bill progress data not yet available._\n")
}
```
:::

:::

---

## Legislature Composition

The Iowa House currently has **`r n_house_R` Republicans** and **`r n_house_D` Democrats**. The Iowa Senate has **`r n_senate_R` Republicans** and **`r n_senate_D` Democrats**.

```{r}
#| layout-ncol: 2
#| fig-height: 3.5

house_parl <- parliament_data(
  election_data = house_comp,
  type          = "semicircle",
  parl_rows     = 6,
  party_seats   = house_comp$seats
)

ggplot(house_parl, aes(x, y, colour = party)) +
  geom_parliament_seats(size = 3.5) +
  scale_colour_manual(
    values = party_colours,
    labels = c("R" = "Republican", "D" = "Democrat")
  ) +
  theme_ggparliament() +
  labs(title = "Iowa House (100 seats)", colour = NULL) +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))

senate_parl <- parliament_data(
  election_data = senate_comp,
  type          = "semicircle",
  parl_rows     = 4,
  party_seats   = senate_comp$seats
)

ggplot(senate_parl, aes(x, y, colour = party)) +
  geom_parliament_seats(size = 5) +
  scale_colour_manual(
    values = party_colours,
    labels = c("R" = "Republican", "D" = "Democrat")
  ) +
  theme_ggparliament() +
  labs(title = "Iowa Senate (50 seats)", colour = NULL) +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))
```

---

## Recent Activity

The table below shows the most recent legislative actions.

```{r}
recent_activity <- history |>
  left_join(bills |> select(bill_id, bill_number, title), by = "bill_id") |>
  arrange(desc(date), desc(sequence)) |>
  head(50) |>
  mutate(
    bill_link = format_bill_link(bill_number, "legislation/")
  ) |>
  select(date, bill_link, chamber, action, title)

gt(recent_activity) |>
  cols_label(
    date = "Date",
    bill_link = "Bill",
    chamber = "Chamber",
    action = "Action",
    title = "Title"
  ) |>
  fmt_markdown(columns = bill_link) |>
  tab_style(
    style = cell_text(whitespace = "nowrap"),
    locations = cells_body(columns = bill_link)
  ) |>
  tab_options(
    table.font.size = px(12),
    data_row.padding = px(4)
  )
```

---

## Browse

- [All Legislation](legislation/index.html) - Browse all bills and resolutions
- [All Legislators](legislators/index.html) - Browse all Representatives and Senators

---

## Data Sources

This site aggregates data from:

- [LegiScan](https://legiscan.com/IA) - Bill text, voting records, and sponsor information
- [Iowa Legislature](https://www.legis.iowa.gov/) - Official state legislature site for lobbyist declarations

---

*Last updated: `r Sys.Date()`*
