---
title: "Legislation - 91st General Assembly"
format:
  html:
    echo: false
    message: false
    warning: false
    toc: true
    toc-location: right
    toc-depth: 2
---

```{r}
library(readr)
library(dplyr)
library(stringr)
library(tidyr)
library(gt)
source(here::here("R/utils.R"))
```

```{r}
# Load data
bills <- read_csv(here::here("legiscan/files_ga91_derived/bills.csv"), show_col_types = FALSE)
people <- read_csv(here::here("legiscan/files_ga91_derived/people.csv"), show_col_types = FALSE)
sponsors <- read_csv(here::here("legiscan/files_ga91_derived/sponsors.csv"), show_col_types = FALSE) |>
  left_join(people)

# Load bill groups (graceful degradation if not yet generated)
bill_groups_path <- here::here("data/bills/groups.rds")
if (file.exists(bill_groups_path)) {
  bill_groups <- read_rds(bill_groups_path)
} else {
  bill_groups <- tibble(
    bill = character(),
    group_status = character(),
    group_size = integer(),
    lead_bill = character()
  )
}
```

## Overview

The 91st Iowa General Assembly has introduced **`r format(nrow(bills), big.mark=",")`** bills and resolutions.

```{r}
# Bill type breakdown
bill_types <- bills |>
  mutate(
    bill_type = case_when(
      str_detect(bill_number, "^HF") ~ "House File (HF)",
      str_detect(bill_number, "^SF") ~ "Senate File (SF)",
      str_detect(bill_number, "^HJR") ~ "House Joint Resolution (HJR)",
      str_detect(bill_number, "^SJR") ~ "Senate Joint Resolution (SJR)",
      str_detect(bill_number, "^HR") ~ "House Resolution (HR)",
      str_detect(bill_number, "^SR") ~ "Senate Resolution (SR)",
      str_detect(bill_number, "^HCR") ~ "House Concurrent Resolution (HCR)",
      str_detect(bill_number, "^SCR") ~ "Senate Concurrent Resolution (SCR)",
      str_detect(bill_number, "^SSB") ~ "Senate Study Bill (SSB)",
      str_detect(bill_number, "^HSB") ~ "House Study Bill (HSB)",
      TRUE ~ "Other"
    )
  ) |>
  count(bill_type, name = "count") |>
  arrange(desc(count))

gt(bill_types) |>
  cols_label(
    bill_type = "Bill Type",
    count = "Count"
  ) |>
  tab_header(title = "Bills by Type")
```

### Status Summary

::: {.grid}

::: {.g-col-6}
```{r}
if (nrow(bill_groups) > 0) {
  status_order <- c("Introduced", "Passed Subcommittee", "Passed Committee",
                    "Passed One Chamber", "Passed Both Chambers", "Signed by Governor")

  status_summary <- bill_groups |>
    count(status, name = "count") |>
    mutate(status = factor(status, levels = status_order)) |>
    arrange(status)

  gt(status_summary) |>
    cols_label(
      status = "Status",
      count = "Count"
    ) |>
    tab_header(title = "Bills by Individual Status")
} else {
  # Fallback to LegiScan status if bill_groups not available
  status_summary <- bills |>
    count(status_desc, name = "count") |>
    arrange(desc(count))

  gt(status_summary) |>
    cols_label(
      status_desc = "Status",
      count = "Count"
    ) |>
    tab_header(title = "Bills by Individual Status")
}
```
:::

::: {.g-col-6}
```{r}
if (nrow(bill_groups) > 0) {
  group_status_order <- c("Introduced", "Passed Subcommittee", "Passed Committee",
                          "Passed One Chamber", "Passed Both Chambers", "Signed by Governor")

  group_status_summary <- bill_groups |>
    count(group_status, name = "count") |>
    mutate(group_status = factor(group_status, levels = group_status_order)) |>
    arrange(group_status)

  gt(group_status_summary) |>
    cols_label(
      group_status = "Group Status",
      count = "Count"
    ) |>
    tab_header(title = "Bills by Group Status") |>
    tab_footnote("Group status reflects the furthest progress of any related bill.")
} else {
  cat("Group status data not yet available. Run `update_bill_groups()` to generate.")
}
```
:::

:::

---

## Recent Bills

Bills with the most recent activity:

```{r}
recent_bills <- bills |>
  arrange(desc(last_action_date)) |>
  head(25) |>
  left_join(
    bill_groups |> select(bill, bill_status = status),
    by = c("bill_number" = "bill")
  ) |>
  mutate(
    bill_link = format_bill_link(bill_number),
    display_status = coalesce(bill_status, status_desc)
  ) |>
  select(bill_link, title, display_status, last_action_date, last_action)

gt(recent_bills) |>
  cols_label(
    bill_link = "Bill",
    title = "Title",
    display_status = "Status",
    last_action_date = "Last Action Date",
    last_action = "Last Action"
  ) |>
  fmt_markdown(columns = bill_link) |>
  tab_style(
    style = cell_text(whitespace = "nowrap"),
    locations = cells_body(columns = bill_link)
  ) |>
  tab_options(
    table.font.size = px(12),
    data_row.padding = px(4)
  )
```

---

## All Legislation

Use the interactive table below to search and filter all bills.

*Group Status shows the furthest progress of any related bill. An asterisk (\*) indicates the group has progressed further than this individual bill.*

```{r}
# Get primary sponsor for each bill
primary_sponsors <- sponsors |>
  filter(position == 1) |>
  select(bill_id, sponsor_name = name)

all_bills <- bills |>
  left_join(primary_sponsors, by = "bill_id") |>
  left_join(
    bill_groups |> select(bill, bill_status = status, group_status, group_size),
    by = c("bill_number" = "bill")
  ) |>
  mutate(
    bill_link = format_bill_link(bill_number),
    chamber = case_when(
      str_detect(bill_number, "^H") ~ "House",
      str_detect(bill_number, "^S") ~ "Senate",
      TRUE ~ "Other"
    ),
    display_status = coalesce(bill_status, status_desc),
    group_status_display = case_when(
      is.na(group_status) ~ NA_character_,
      group_size == 1 ~ group_status,
      group_status == display_status ~ group_status,
      TRUE ~ paste0(group_status, "*")
    )
  ) |>
  arrange(desc(last_action_date)) |>
  select(bill_link, title, chamber, sponsor_name, display_status, group_status_display, last_action_date)

gt(all_bills) |>
  cols_label(
    bill_link = "Bill",
    title = "Title",
    chamber = "Chamber",
    sponsor_name = "Primary Sponsor",
    display_status = "Status",
    group_status_display = "Group Status",
    last_action_date = "Last Action"
  ) |>
  fmt_markdown(columns = bill_link) |>
  tab_style(
    style = cell_text(whitespace = "nowrap"),
    locations = cells_body(columns = bill_link)
  ) |>
  tab_options(
    table.font.size = px(12),
    data_row.padding = px(4)
  ) |>
  opt_interactive(
    use_search = TRUE,
    use_filters = TRUE,
    use_page_size_select = TRUE,
    page_size_default = 25
  )
```

---

## House Files

```{r}
house_files <- bills |>
  filter(str_detect(bill_number, "^HF")) |>
  left_join(primary_sponsors, by = "bill_id") |>
  left_join(
    bill_groups |> select(bill, bill_status = status, group_status, group_size),
    by = c("bill_number" = "bill")
  ) |>
  mutate(
    bill_link = format_bill_link(bill_number),
    display_status = coalesce(bill_status, status_desc),
    group_status_display = case_when(
      is.na(group_status) ~ NA_character_,
      group_size == 1 ~ group_status,
      group_status == display_status ~ group_status,
      TRUE ~ paste0(group_status, "*")
    )
  ) |>
  arrange(as.numeric(str_extract(bill_number, "\\d+"))) |>
  select(bill_link, title, sponsor_name, display_status, group_status_display, last_action_date)

gt(house_files) |>
  cols_label(
    bill_link = "Bill",
    title = "Title",
    sponsor_name = "Primary Sponsor",
    display_status = "Status",
    group_status_display = "Group Status",
    last_action_date = "Last Action"
  ) |>
  fmt_markdown(columns = bill_link) |>
  tab_style(
    style = cell_text(whitespace = "nowrap"),
    locations = cells_body(columns = bill_link)
  ) |>
  tab_options(
    table.font.size = px(12),
    data_row.padding = px(4)
  ) |>
  opt_interactive(
    use_search = TRUE,
    use_filters = TRUE,
    use_page_size_select = TRUE,
    page_size_default = 25
  )
```

---

## Senate Files

```{r}
senate_files <- bills |>
  filter(str_detect(bill_number, "^SF")) |>
  left_join(primary_sponsors, by = "bill_id") |>
  left_join(
    bill_groups |> select(bill, bill_status = status, group_status, group_size),
    by = c("bill_number" = "bill")
  ) |>
  mutate(
    bill_link = format_bill_link(bill_number),
    display_status = coalesce(bill_status, status_desc),
    group_status_display = case_when(
      is.na(group_status) ~ NA_character_,
      group_size == 1 ~ group_status,
      group_status == display_status ~ group_status,
      TRUE ~ paste0(group_status, "*")
    )
  ) |>
  arrange(as.numeric(str_extract(bill_number, "\\d+"))) |>
  select(bill_link, title, sponsor_name, display_status, group_status_display, last_action_date)

gt(senate_files) |>
  cols_label(
    bill_link = "Bill",
    title = "Title",
    sponsor_name = "Primary Sponsor",
    display_status = "Status",
    group_status_display = "Group Status",
    last_action_date = "Last Action"
  ) |>
  fmt_markdown(columns = bill_link) |>
  tab_style(
    style = cell_text(whitespace = "nowrap"),
    locations = cells_body(columns = bill_link)
  ) |>
  tab_options(
    table.font.size = px(12),
    data_row.padding = px(4)
  ) |>
  opt_interactive(
    use_search = TRUE,
    use_filters = TRUE,
    use_page_size_select = TRUE,
    page_size_default = 25
  )
```

---

[Back to Home](../index.html)

*Last updated: `r Sys.Date()`*
