---
# title: "bill_template"
format: 
  html:
    echo: false
    message: false
    warning: false
    toc: true
    toc-location: right
    toc-depth: 2
    
params:
  bill_num: "HF856"
---

```{r}
library(jsonlite)
library(readr)

library(dplyr)
library(purrr)
library(stringr)
library(tidyr)

library(gt)
library(sf)
library(leaflet)
library(htmltools)

source(here::here("R/utils.R"))

```


```{r}
bill <- read_json(here::here(str_glue("legiscan/files_ga91_json/bill/{params$bill_num}.json"))) |>
  pluck("bill")

people <- read_csv(here::here(str_glue("legiscan/files_ga91/people.csv")), show_col_types = FALSE)
bills_csv <- read_csv(here::here("legiscan/files_ga91/bills.csv"), show_col_types = FALSE)

sponsors <- bill |>
  pluck("sponsors") |>
  items_to_df()

history <- bill |>
  pluck("history") |>
  items_to_df()


# Votes lists more than just the committee vote for HF1
votes <- bill |>
  pluck("votes") |>
  items_to_df()

calendar <- bill |>
  pluck("calendar") |>
  items_to_df()

# Load processed vote summaries for party breakdown and classification
vote_summaries <- read_rds(here::here("data/vote_summaries_91st_ga.rds"))
bill_vote_summaries <- vote_summaries |>
  filter(bill_id == bill$bill_id)

# Load bill groups (graceful degradation if not yet generated)
bill_groups_path <- here::here("data/bill_groups.rds")
if (file.exists(bill_groups_path)) {
  bill_groups <- read_rds(bill_groups_path)
  bill_group_info <- bill_groups |> filter(bill == params$bill_num)
} else {
  bill_groups <- tibble(
    bill = character(),
    status = character(),
    group_status = character(),
    group_size = integer(),
    lead_bill = character(),
    group_members = list()
  )
  bill_group_info <- tibble(
    bill = params$bill_num,
    status = NA_character_,
    group_status = NA_character_,
    group_size = 1L,
    lead_bill = NA_character_,
    group_members = list(params$bill_num)
  )
}

# Load district shapefiles for vote maps
senate_districts <- st_read(here::here("data/Districts/Plan2_Senate.shp"), quiet = TRUE) |>
  st_transform(crs = 4326) |>
  transmute(District = as.numeric(DISTRICT))

house_districts <- st_read(here::here("data/Districts/Plan2_House.shp"), quiet = TRUE) |>
  st_transform(crs = 4326) |>
  transmute(District = as.numeric(DISTRICT))

# Scrape lobbyist declarations
lobby <- tryCatch(
  scrape_lobbyist_declarations(params$bill_num),
  error = function(e) {
    message(paste("Lobbyist scrape failed for", params$bill_num, ":", e$message))
    NULL
  }
)

lobby_for <- if (!is.null(lobby) && nrow(lobby) > 0) sum(lobby$most_recent_declaration == "For", na.rm = TRUE) else 0
lobby_against <- if (!is.null(lobby) && nrow(lobby) > 0) sum(lobby$most_recent_declaration == "Against", na.rm = TRUE) else 0
lobby_undecided <- if (!is.null(lobby) && nrow(lobby) > 0) sum(lobby$most_recent_declaration == "Undecided", na.rm = TRUE) else 0
```


# [`r params$bill_num`](`r bill$state_link`){target="_blank"}

`r bill$title`

```{r}
#| output: asis
if (nrow(bill_group_info) > 0 && !is.na(bill_group_info$status)) {
  cat(paste0("**Status:** ", bill_group_info$status, "\n\n"))

  if (bill_group_info$group_size > 1 && bill_group_info$status != bill_group_info$group_status) {
    lead_bill_display <- str_replace(bill_group_info$lead_bill, "^([A-Z]+)(\\d+)$", "\\1 \\2")
    cat(paste0("**Group Status:** ", bill_group_info$group_status,
               " (via [", lead_bill_display, "](", bill_group_info$lead_bill, ".html))\n\n"))
  }
} else {
  cat("**Status:** ", bill$status, "\n\n")
}

cat(paste0("*Most Recent Action: ", if(!is.null(history) && nrow(history) > 0) history$action[1] else "None", "*\n\n"))

if (!is.null(lobby) && nrow(lobby) > 0) {
  cat(paste0("**Lobbyist Declarations:** ", lobby_for, " For | ", lobby_against, " Against | ", lobby_undecided, " Undecided\n\n"))
}
```

---

## Sponsors
```{r}
if (!is.null(sponsors) && nrow(sponsors) > 0) {
  gt(sponsors |> select(role, name, party, district)) |>
    cols_label(role = "Role", name = "Name", party = "Party", district = "District")
} else {
  cat("No sponsors listed.")
}
```

---

## Related Legislation

```{r}
#| output: asis
if (nrow(bill_group_info) > 0 && bill_group_info$group_size > 1) {
  cat(paste0("This bill is part of a group of **", bill_group_info$group_size, "** related bills. "))

  if (bill_group_info$status != bill_group_info$group_status) {
    lead_bill_display <- str_replace(bill_group_info$lead_bill, "^([A-Z]+)(\\d+)$", "\\1 \\2")
    cat(paste0("The group's furthest progress is **", bill_group_info$group_status,
               "**, achieved by [", lead_bill_display, "](", bill_group_info$lead_bill, ".html).\n\n"))
  } else {
    cat("\n\n")
  }
} else {
  cat("No related bills identified.\n")
}
```

```{r}
if (nrow(bill_group_info) > 0 && bill_group_info$group_size > 1) {
  related_bills_list <- unlist(bill_group_info$group_members)

  related_df <- bill_groups |>
    filter(bill %in% related_bills_list, bill != params$bill_num) |>
    left_join(
      bills_csv |> select(bill_number, title),
      by = c("bill" = "bill_number")
    ) |>
    mutate(bill_link = format_bill_link(bill, "")) |>
    arrange(desc(status_level)) |>
    select(bill_link, title, status)

  if (nrow(related_df) > 0) {
    gt(related_df) |>
      cols_label(
        bill_link = "Bill",
        title = "Title",
        status = "Status"
      ) |>
      fmt_markdown(columns = bill_link) |>
      tab_options(
        table.font.size = px(12),
        data_row.padding = px(4)
      )
  }
}
```

---

::: {.panel-tabset}

## Legislation History
```{r}
if (!is.null(history) && nrow(history) > 0) {
  gt(history |>
    mutate(chamber = case_when(chamber == "H" ~ "House", chamber == "S" ~ "Senate", .default = chamber)) |>
    select(date, chamber, action)) |>
    cols_label(date = "Date", chamber = "Chamber", action = "Action")
} else {
  cat("No legislative history recorded yet.")
}
```


## Votes

### Vote Summaries
```{r}
if (!is.null(votes) && nrow(votes) > 0) {
  votes_filtered <- votes |> filter(!str_detect(desc, "Report$"))
  if (nrow(votes_filtered) > 0) {
    # Merge with processed vote data for party breakdown
    votes_display <- votes_filtered |>
      left_join(
        bill_vote_summaries |> select(
          roll_call_id, vote_dem_yes, vote_dem_no, vote_gop_yes, vote_gop_no, vote_outcome
        ),
        by = "roll_call_id"
      )

    gt(votes_display |> select(
      date, chamber, desc, yea, nay, nv, absent,
      vote_dem_yes, vote_dem_no, vote_gop_yes, vote_gop_no, vote_outcome
    )) |>
      cols_label(
        date = "Date", chamber = "Chamber", desc = "Description",
        yea = "Yea", nay = "Nay", nv = "NV", absent = "Absent",
        vote_dem_yes = "D Yea", vote_dem_no = "D Nay",
        vote_gop_yes = "R Yea", vote_gop_no = "R Nay",
        vote_outcome = "Classification"
      ) |>
      tab_spanner(label = "Total", columns = c(yea, nay, nv, absent)) |>
      tab_spanner(label = "By Party", columns = c(vote_dem_yes, vote_dem_no, vote_gop_yes, vote_gop_no))
  } else {
    cat("No floor votes recorded yet.")
  }
} else {
  cat("No votes recorded yet.")
}
```

### Individual Votes

::: {.panel-tabset}

```{r}
#| results: "asis"

if (!is.null(votes) && nrow(votes) > 0) {
  votes_filtered <- votes |> filter(!str_detect(desc, "Report$"))

  vote_child_template <- c(
    "## `r vote_row$chamber` : `r vote_row$date`",
    "",
    "#### `r vote_row$desc`",
    "",
    "```{r}",
    "#| echo: false",
    "#| message: false",
    "#| warning: false",
    "vote <- read_json(here::here(paste0('legiscan/files_ga91_json/vote/', vote_num, '.json'))) |>",
    "  pluck('roll_call') |>",
    "  pluck('votes') |>",
    "  items_to_df() |>",
    "  left_join(people |> select(people_id, name, party, district)) |>",
    "  select(-people_id, -vote_id) |>",
    "  arrange(district)",
    "",
    "summary_tbl <- vote |>",
    "  count(party, vote_text) |>",
    "  mutate(vote_text = factor(vote_text, levels = c('Yea', 'Nay', 'Absent'))) |>",
    "  arrange(vote_text, party) |>",
    "  pivot_wider(names_from = vote_text, values_from = n) |>",
    "  mutate(across(everything(), .fns = function(x){replace_na(x, 0)}))",
    "",
    "gt(summary_tbl, rowname_col = 'party') |>",
    "  tab_stubhead(label = 'Party')",
    "```",
    "",
    "```{r}",
    "#| echo: false",
    "#| message: false",
    "#| warning: false",
    "vote_chamber <- vote_row$chamber",
    "districts_sf <- if (vote_chamber == 'H') house_districts else senate_districts",
    "",
    "vote_map_data <- vote |>",
    "  mutate(",
    "    district_num = parse_district_number(district),",
    "    vote_category = case_when(",
    "      party == 'D' & vote_text == 'Yea' ~ 'Democrat: Yea',",
    "      party == 'D' & vote_text == 'Nay' ~ 'Democrat: Nay',",
    "      party == 'R' & vote_text == 'Yea' ~ 'Republican: Yea',",
    "      party == 'R' & vote_text == 'Nay' ~ 'Republican: Nay',",
    "      .default = 'Absent/NV'",
    "    ),",
    "    fill_color = case_when(",
    "      vote_category == 'Democrat: Yea' ~ '#2320E6',",
    "      vote_category == 'Democrat: Nay' ~ '#A4A3F6',",
    "      vote_category == 'Republican: Yea' ~ '#DC0309',",
    "      vote_category == 'Republican: Nay' ~ '#EE8181',",
    "      .default = '#CCCCCC'",
    "    )",
    "  )",
    "",
    "map_sf <- districts_sf |>",
    "  left_join(vote_map_data |> select(district_num, name, party, vote_text, vote_category, fill_color),",
    "            by = c('District' = 'district_num'))",
    "",
    "leaflet(map_sf) |>",
    "  addProviderTiles(providers$CartoDB.Positron) |>",
    "  addPolygons(",
    "    fillColor = ~ifelse(is.na(fill_color), '#CCCCCC', fill_color),",
    "    fillOpacity = 0.7,",
    "    color = '#333333',",
    "    weight = 1,",
    "    label = ~paste0(ifelse(is.na(name), paste('District', District), name),",
    "                    ' (', ifelse(is.na(party), '?', party), '-', District, '): ',",
    "                    ifelse(is.na(vote_text), 'No vote', vote_text))",
    "  ) |>",
    "  addLegend(",
    "    position = 'bottomright',",
    "    colors = c('#2320E6', '#A4A3F6', '#DC0309', '#EE8181', '#CCCCCC'),",
    "    labels = c('Democrat: Yea', 'Democrat: Nay', 'Republican: Yea', 'Republican: Nay', 'Absent/NV'),",
    "    title = 'Vote'",
    "  )",
    "```",
    "",
    "```{r}",
    "#| echo: false",
    "#| message: false",
    "#| warning: false",
    "gt(vote)",
    "```",
    ""
  )

  res <- lapply(votes_filtered |> pull(roll_call_id), function(vote_num) {
    vote_row <- votes |> filter(roll_call_id == vote_num)
    knitr::knit_child(text = vote_child_template, envir = environment(), quiet = TRUE)
  })

  cat(unlist(res), sep = "\n")
}
```


:::

## Lobbyist Declarations

```{r}
#| output: asis
if (!is.null(lobby) && nrow(lobby) > 0) {
  cat(paste0("**For:** ", lobby_for, " | **Against:** ", lobby_against, " | **Undecided:** ", lobby_undecided, "\n\n"))
} else {
  cat("Lobbyist declaration data is currently unavailable for this bill.\n")
}
```

```{r}
if (!is.null(lobby) && nrow(lobby) > 0) {
  gt(lobby |>
    select(client, most_recent_declaration, date) |>
    arrange(most_recent_declaration, client)) |>
    cols_label(
      client = "Client",
      most_recent_declaration = "Declaration",
      date = "Last Updated"
    ) |>
    fmt_date(columns = date, date_style = "yMMMd") |>
    tab_options(
      table.font.size = px(12),
      data_row.padding = px(4)
    )
}
```


## Calendar Items
```{r}
if (!is.null(calendar) && nrow(calendar) > 0) {
  gt(calendar |> select(date, time, type, location, description)) |>
    cols_label_with(fn = stringr::str_to_title)
} else {
  cat("No calendar items scheduled.")
}
```

:::



