---
pagetitle: "`r jsonlite::read_json(here::here(paste0('legiscan/files_ga91_json/people/', params$people_id, '.json')))$person$name`"
format: 
  html:
    echo: false
    message: false
    warning: false
    toc: true
    toc-location: right
    toc-depth: 2
    
params:
  people_id: 19548
---

```{r}
library(jsonlite)
library(readr)
library(dplyr)
library(purrr)
library(stringr)
library(tidyr)
library(gt)
library(sf)
library(leaflet)
library(htmltools)
source(here::here("R/utils.R"))
```

```{r}
items_to_df <- function(list){
  map(list, as.data.frame) |>
    bind_rows()
}
```

```{r}
# Read legislator data
legislator <- read_json(here::here(str_glue("legiscan/files_ga91_json/people/{params$people_id}.json"))) |>
  pluck("person")

# Read related data
bills <- read_csv(here::here("legiscan/files_ga91_derived/bills.csv"), show_col_types = FALSE)
sponsors <- read_csv(here::here("legiscan/files_ga91_derived/sponsors.csv"), show_col_types = FALSE)
votes <- read_csv(here::here("legiscan/files_ga91_derived/votes.csv"), show_col_types = FALSE)
rollcalls <- read_csv(here::here("legiscan/files_ga91_derived/rollcalls.csv"), show_col_types = FALSE)

votes <- left_join(votes, rollcalls, by = join_by(roll_call_id)) |>
  left_join(bills |> select(bill_id, bill_number, title), by = join_by(bill_id))

# Filter to this legislator's activity
leg_sponsored_bills <- sponsors |>
  mutate(sponsor_num = n(), .by = bill_id) |>
  filter(people_id == params$people_id) |>
  left_join(bills, by = "bill_id") |>
  arrange(desc(status_date))

leg_votes <- votes |>
  filter(
    people_id == params$people_id,
    !str_detect(description, "Report$")
  )

# Load processed vote data for party alignment analysis
legislator_stats <- read_rds(here::here("data/legislator_voting_stats_91st_ga.rds"))
vote_records <- read_rds(here::here("data/vote_records_91st_ga.rds"))

leg_stats <- legislator_stats |> filter(people_id == params$people_id)
leg_vote_records <- vote_records |> filter(people_id == params$people_id)

# Load simplified district geometries and join legislator names
people <- read_csv(here::here("legiscan/files_ga91_derived/people.csv"), show_col_types = FALSE) |>
  filter(party_id != 0) |>
  mutate(district_num = parse_district_number(district)) |>
  select(name, party, district, district_num, role)

senate_districts <- read_rds(here::here("data/senate_districts.rds")) |>
  left_join(people |> filter(role == "Sen"), by = c("District" = "district_num"))

house_districts <- read_rds(here::here("data/house_districts.rds")) |>
  left_join(people |> filter(role == "Rep"), by = c("District" = "district_num"))
```

# `r legislator$name` (`r legislator$party`-`r legislator$district`)

**Role:** `r legislator$role`  
**Party:** `r legislator$party`  
**District:** `r legislator$district`  
**State:** Iowa

---

**External Links:**

```{r}
#| output: asis
leg_email <- legislator$bio$social$email
leg_website <- legislator$bio$links$official$website

if (!is.null(leg_website) && nchar(leg_website) > 0) {
  cat(sprintf("- [Iowa Legislature Page](%s){target=\"_blank\"}\n", leg_website))
}
if (!is.null(leg_email) && nchar(leg_email) > 0) {
  cat(sprintf("- [%s](mailto:%s)\n", leg_email, leg_email))
}
cat(sprintf("- [Ballotpedia Profile](https://ballotpedia.org/%s){target=\"_blank\"}\n", legislator$ballotpedia))
cat(sprintf("- [VoteSmart Profile](https://justfacts.votesmart.org/candidate/%s){target=\"_blank\"}\n", legislator$votesmart_id))
```

### District Map

```{r}
district_num <- parse_district_number(legislator$district)
districts_sf <- if (legislator$role == "Sen") senate_districts else house_districts

highlight <- districts_sf |> filter(District == district_num)
others <- districts_sf |> filter(District != district_num)

party_color <- if (legislator$party == "D") "#2320E6" else "#DC0309"

bbox <- st_bbox(highlight)

leaflet() |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addPolygons(
    data = others,
    fillColor = "#CCCCCC",
    fillOpacity = 0.2,
    color = "#000000",
    weight = 1,
    label = ~ifelse(!is.na(name), paste0(name, " (", party, "-", district, ")"), paste("District", District))
  ) |>
  addPolygons(
    data = highlight,
    fillColor = party_color,
    fillOpacity = 0.6,
    color = party_color,
    weight = 3,
    label = ~paste(legislator$name, "-", legislator$district)
  ) |>
  fitBounds(bbox[["xmin"]], bbox[["ymin"]], bbox[["xmax"]], bbox[["ymax"]])
```

---

::: {.panel-tabset}

## Legislative Summary

```{r}
# Calculate summary statistics
total_sponsored <- nrow(leg_sponsored_bills)
total_votes <- nrow(leg_votes)

# Load bill groups for outcome context
bill_groups <- read_rds(here::here("data/bill_groups.rds"))

# Sponsorship with outcomes
if(nrow(leg_sponsored_bills) > 0) {
  sponsor_summary <- leg_sponsored_bills |>
    mutate(
      Role = case_when(
        position == 1 & sponsor_num == 1 ~ "Sole Sponsor",
        position == 1 ~ "Primary Sponsor",
        position >= 2 ~ "Co-Sponsor"
      )
    ) |>
    left_join(
      bill_groups |> select(bill, group_status, group_size),
      by = c("bill_number" = "bill")
    ) |>
    mutate(group_status = coalesce(group_status, status_desc))
} else {
  sponsor_summary <- data.frame()
}
```

### Sponsorship Activity

**Total Bills Sponsored:** `r total_sponsored`

```{r}
if(nrow(sponsor_summary) > 0) {
  # Bill outcome by sponsorship role
  status_order <- c("Introduced", "Passed Subcommittee", "Passed Committee",
                     "Passed One Chamber", "Passed Both Chambers", "Signed by Governor")

  sponsor_outcome <- sponsor_summary |>
    mutate(group_status = factor(group_status, levels = status_order)) |>
    count(Role, group_status, .drop = FALSE) |>
    pivot_wider(names_from = group_status, values_from = n, values_fill = 0) |>
    filter(rowSums(pick(-Role)) > 0) |>
    mutate(Total = rowSums(pick(-Role))) |>
    arrange(desc(Total))

  gt(sponsor_outcome) |>
    cols_label(Role = "Sponsorship Role") |>
    tab_spanner(label = "Bill Group Outcome", columns = c(-Role, -Total))
} else {
  cat("No sponsorship data available.\n")
}
```


### Party Alignment

```{r}
if(nrow(leg_stats) > 0 && leg_stats$total_votes > 0) {
  party_alignment <- data.frame(
    Metric = c(
      "Votes With Party",
      "Votes Against Party",
      "Party Alignment Rate",
      "Participation Rate"
    ),
    Value = c(
      as.character(leg_stats$votes_with_party),
      as.character(leg_stats$votes_against_party),
      sprintf("%.1f%%", leg_stats$with_party_pct * 100),
      sprintf("%.1f%%", leg_stats$participation_rate * 100)
    )
  )

  gt(party_alignment) |>
    cols_label(Metric = "", Value = "")
} else {
  cat("Insufficient voting data for party alignment analysis.\n")
}
```

<!-- ### Votes by Category

```{r}
#| eval: false
if(nrow(leg_vote_records) > 0) {
  category_breakdown <- leg_vote_records |>
    filter(vote_text %in% c("Yea", "Nay")) |>
    count(vote_outcome, vote_text) |>
    pivot_wider(names_from = vote_outcome, values_from = n, values_fill = 0) |>
    rename(Vote = vote_text)

  gt(category_breakdown) |>
    tab_spanner(label = "Vote Category", columns = -Vote)
} else {
  cat("Insufficient voting data for category breakdown.\n")
}
```
-->

## Sponsored Bills

```{r}
if(nrow(leg_sponsored_bills) > 0) {
  leg_sponsored_bills |>
    mutate(bill_link = format_bill_link(bill_number, "../legislation/")) |>
    select(bill_link, title, status_desc, status_date, position, sponsor_num) |>
    mutate(
      Sponsor_Role = case_when(
        position == 1 & sponsor_num == 1 ~ "Sole Sponsor",
        position == 1 ~ "Primary Sponsor",
        position >= 2 ~ "Co-Sponsor"
      )
    ) |>
    select(-position) |>
    relocate(sponsor_num, .after = Sponsor_Role) |>
    gt() |>
    cols_label(
      bill_link = "Bill",
      title = "Title",
      status_desc = "Status",
      status_date = "Status Date",
      Sponsor_Role = "Role",
      sponsor_num = "# of Sponsors"
    ) |>
    fmt_markdown(columns = bill_link) |>
    tab_style(
      style = cell_text(whitespace = "nowrap"),
      locations = cells_body(columns = bill_link)
    ) |>
    tab_options(
      table.font.size = px(12)
    )
} else {
  cat("No sponsored bills found.\n")
}
```

## Voting Record

```{r}
if(nrow(leg_votes) > 0) {
  gt(leg_votes |>
    mutate(
      bill_link = format_bill_link(bill_number, "../legislation/"),
      description = simplify_vote_desc(description)
    ) |>
    select(bill_link, title, date, description, vote_desc)) |>
    cols_label(
      bill_link = "File #",
      title = "Title",
      date = "Vote Date",
      description = "Vote Type",
      vote_desc = "Vote"
    ) |>
    fmt_markdown(columns = bill_link) |>
    tab_style(
      style = cell_text(whitespace = "nowrap"),
      locations = cells_body(columns = bill_link)
    )
} else {
  cat("No voting record found.\n")
}
```

## Vote Analysis

### Votes by Classification

```{r}
if(nrow(leg_vote_records) > 0) {
  outcome_breakdown <- leg_vote_records |>
    filter(vote_text %in% c("Yea", "Nay")) |>
    group_by(vote_outcome) |>
    summarize(
      count = n(),
      with_party = sum(vote_with_party == TRUE, na.rm = TRUE),
      against_party = sum(vote_with_party == FALSE, na.rm = TRUE),
      .groups = "drop"
    ) |>
    mutate(pct_of_total = count / sum(count) * 100) |>
    arrange(desc(count))

  gt(outcome_breakdown) |>
    cols_label(
      vote_outcome = "Vote Type",
      count = "Total",
      with_party = "With Party",
      against_party = "Against Party",
      pct_of_total = "% of Votes"
    ) |>
    fmt_number(columns = pct_of_total, decimals = 1)
} else {
  cat("Insufficient voting data for analysis.\n")
}
```

### Votes Against Party

```{r}
if(nrow(leg_vote_records) > 0) {
  against_party <- leg_vote_records |>
    filter(vote_with_party == FALSE) |>
    left_join(bills |> select(bill_id, title), by = "bill_id") |>
    mutate(
      bill_link = format_bill_link(bill_number, "../legislation/"),
      vote_desc = simplify_vote_desc(vote_desc)
    ) |>
    select(bill_link, title, vote_date, vote_desc, vote_text, party_position, vote_outcome) |>
    arrange(desc(vote_date))

  if(nrow(against_party) > 0) {
    gt(against_party) |>
      cols_label(
        bill_link = "Bill",
        title = "Title",
        vote_date = "Date",
        vote_desc = "Description",
        vote_text = "Vote",
        party_position = "Party Position",
        vote_outcome = "Classification"
      ) |>
      fmt_markdown(columns = bill_link) |>
      tab_style(
        style = cell_text(whitespace = "nowrap"),
        locations = cells_body(columns = bill_link)
      )
  } else {
    cat("This legislator has not voted against their party.\n")
  }
} else {
  cat("Insufficient voting data for analysis.\n")
}
```

:::

---

## Notes

This profile was generated from LegiScan data for the Iowa Legislature.  
Last updated: `r Sys.Date()`